---
title: "Defining non-standard employment"
description: |
  Lecture 1
# author:
#   - name: Kostiantyn Ovsiannikov
date: 2021-11-01
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    self_contained: false
    # code_folding: hide
creative_commons: CC BY
---


```{r setup, include=FALSE}
library(knitr)
library(summarytools)
library(here)
library(ymlthis)
library(tidyverse)
library(fontawesome)
library(readr)
library(ggthemes)
library(ggpubr)
library(RColorBrewer)
library(kableExtra)
library(data.table)
library(metathis)
library(xaringanthemer)
library(xaringanExtra)
library(data.table)
library(RefManageR)
library(plm)
library(mgcv)
library(mgcViz)
library(stargazer)
library(plotly)
library(htmlwidgets)
library(modelsummary)
```


```{r echo=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(collapse = TRUE,
                      fig.retina = 3)
```


```{r, load_refs, include=FALSE, cache=FALSE}
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "alphabetic",
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib(here("_posts", "lecture-1", "Labor-Dualism.bib"), check = FALSE)
```

```{r echo=FALSE, include=FALSE}
capture.output(print(myBib["kalleberg2000ars"]))
```

<!-- ## Article discussed -->

::: {.define .article style="color: brown;"}
[Kalleberg, A. L. (2000). Nonstandard employment relations: Part-time, temporary and contract work. *Annual review of sociology, 26* (1), 341-365](https://www.annualreviews.org/doi/abs/10.1146/annurev.soc.26.1.341)
:::


## Background

1970s' **oil crises** `r fa("arrow-alt-circle-right", fill = "#cc0033")` sluggish economic growth `r fa("arrow-alt-circle-right", fill = "#cc0033")` high unemployment in the West `r fa("arrow-alt-circle-right", fill = "#cc0033")` inability to generate full employment for everyone.

Nonstandard employment relations are not new.
In fact, internal labor markets in the post-World War II period were more of an **exception** than a regularity.
The main organizational concern of that time: how to internalize their workforce `r fa("arrow-alt-circle-right", fill = "#cc0033")` develop their skills; protect from external market competition.

Main organizational concern today: achieving flexibility by **externalizing** some of their activites.


## Part-time work

### Definitions, incidence, trends

Part-time work -- regular wage employment in which the hours of work are **less than "normal"**.
Different definitions across countries:

- <span style="color:darkred">U.S.</span>: < 35 hours/week
- <span style="color:darkred">Japan</span>: related to a **status within the firm** -- not to hours worked `r fa("arrow-alt-circle-right", fill = "#cc0033")` around 30\% of "part-timers" work full time

Part time proportions:

 - U.S.: 20\%
 - Netherlands: 38\%
 - Scandinavian countries: 20\%
 - Italy: 5\%

<br>

Rapid increase of part-time employment in Europe:

 - a way to alleviate unemployment
 - major source of employment growth since 1980s
 
<br>
 
In all industrial countries, most part-time workers are **women**:

 - U.S.: 65\%
 - Germany, France: 90\%
 - Japan: 80\%

Lower proportion in the U.S. -- many students (both male and female) work part-time.


### Demand for part-time jobs

- Pre-1970s U.S. -- growth of **voluntary** part-time work `r fa("arrows-alt-h", fill = "blue")` 
The need of **workforce** for shorter hours.
- Post-1970s U.S. -- growth of **involuntary** part-time work `r fa("arrows-alt-h", fill = "blue")`
**Employers'** preferences for lower costs and flexible staffing.
Expansion of industries such as: services, retail trade, finance, insurance, real estate.


### Good vs. bad part-time jobs

Europe: part-time work is often associated with low-paid, low-status jobs (sales, catering, cleaning). 
Legal differences:

 - Sweden, France, Belgium, Netherlands, Spain: **equal treatment** between full- and part-time workers
 - U.K., Germany, Japan: part-timers are excluded from comprehensive social coverage
 

## Employment intermediaries

### Temporary help agencies

"Detached workers": **triangular** employment relations where a worker establishes connections with several employers.

Changes in temporary work `r fa("arrows-alt-h", fill = "blue")` business cycles, economic fluctuations.
Temporary agencies were largely banned before 1980s.
About 60\% work in temporary jobs involuntarily.
Temporary help agencies -- "reserve labor army" that helps employers to address under-staffing and over-staffing. 

Major temp agencies `r fa("arrows-alt-h", fill = "blue")` firms.
Use of temporary workers has become a **permanent strategy** for them.
<span style="color:darkred">Yet</span>, hourly costs to employees may be more for temporaries than for permanent employees.
In 1995, client firms paid 40\% more for temporary workers than they received wages.
However, these wages are relatively low.

Deliberate **wage differences** -- illegal vis-à-vis full-time regular workers, but doable in case of full-time vs. temporary workers. 
More fringe benefits for regular employees `r fa("arrow-alt-circle-right", fill = "#cc0033")` higher likelihood to use temporary help agencies.

Triangular employment relationship: 

 - legal employer: <span style="color:darkred">temporary agency</span>
 - employee supervisor: <span style="color:darkred">client organization</span>
 
Main issues: liability for accidents and other aspects `r fa("arrow-alt-circle-right", fill = "#cc0033")`
Firms tend to avoid being a **joint employer** by specifying a period, after which a temp is obliged to leave a company.

Problems of control `r fa("arrow-alt-circle-right", fill = "#cc0033")` difficulties with unionizing temps.
If a client firm is not a "joint employer" `r fa("arrow-alt-circle-right", fill = "#cc0033")` it can fire employees for joining a union.
Presence of temporary agency `r fa("arrow-alt-circle-right", fill = "#cc0033")` **decreased** mobility among low-rank permanent workers, **increased** mobility among advanced workers.

Further careers:

 - about 58\% find permanent jobs after a year-and-half
 - about 42\% either become unemployed or remain with an agency
 
Temps: 1/3 less salary and much fewer fringe benefits than regulars.
Largest wage differences -- for clerical and industrial workers.
<span style="color:darkred">However</span>, temp **engineers** and technicians often earn more than their regular counterparts.


### Contract work

Contrary to temp agencies, contract companies supervise their employees' work.
It has existed for a long time in the **construction industry**.

Since 1970s -- firms started to contract out (outsource) services previously done in-house:

 - advertising
 - credit
 - mailing
 - maintenance & cleaning
 - data processing
 - R&D
 
Cost-saving strategy: firms have **little incentives** of paying high wages to replaceable, easily-monitored workers who perform **peripheral duties**.

<span style="color:darkred">Problem of control</span>: managers sometimes avoid training contract workers to avoid **co-employment liabilities**.

Presence of contractors in some industries (e.g. aerospace) `r fa("arrow-alt-circle-right", fill = "#cc0033")` less employee trust.
Yet, contractors can earn more than regulars and usually perform not worse.


## Short-term employment

### Types, incidence, trends

EU -- 25\% increase during 1990s, but representing only 11.5\% of the workforce.
Has grown less than part-time work.
Unlike part-time work, does not normally mitigate unemployment `r fa("arrow-alt-circle-right", fill = "#cc0033")` indicator of a **weak worker position** in a labor market.
Particularly important for countries with strong anti-layoff regulations.
Labor market rigidities `r fa("arrows-alt-h", fill = "blue")` greater use of temporary workers:

 - <span style="color:darkred">Spain, France</span> (**strong restrictions** on dismissals): proportion of fixed-term contracts between 1985 and 1996 -- **doubled up** to 34\% and 13\% respectively
 - <span style="color:darkred">U.K., Germany</span>: proportion of fixed-term contracts between 1985 and 1996 -- stood at 7\% and 11\% respectively
 - <span style="color:darkred">Belgium, Greece, Portugal</span>: **decrease** in fixed-term contracts' proportion between 1985 and 1996
 
 
### Organizations' use of short-term temporaries

U.K.: older organizations are less likely to use fixed-term temporaries.
Temporaries' productivity is usually **lower** than regulars'.
Use of temporaries `r fa("arrow-alt-circle-right", fill = "#cc0033")` waste of organizational training efforts.
Inter-country differences:

 - <span style="color:darkred">U.K.</span>: obtaining **numerical flexibility** (ability to decrease workforce size)
 - <span style="color:darkred">Germany</span>: obtaining **functional flexibility** (employees' ability to do variety of tasks); necessary skills' acquisition `r fa("arrow-alt-circle-right", fill = "#cc0033")` possible permanent contract
 

### Different types of "flexibility" (`r Citep(myBib, "atkinson1984manpower", .opts = list(cite.style = "authoryear", style = "html"))`)

- <span style="color:darkred">numerical</span>: loosening the hire/fire procedure `r fa("arrow-alt-circle-right", fill = "#cc0033")` promptly adjusting the number of workers to employer's needs

- <span style="color:darkred">functional</span>: smooth redeployment of workers between activities and tasks `r fa("arrow-alt-circle-right", fill = "#cc0033")` anticipating the technological and production changes

- <span style="color:darkred">financial</span>: widening payment differentials between skilled and unskilled workers `r fa("arrow-alt-circle-right", fill = "#cc0033")` assessment-based system facilitating numerical and functional flexibility
 

## Contingent work

Related to short-term, unstable employment.
Conditional on employers' need for labor `r fa("arrow-alt-circle-right", fill = "#cc0033")` weak worker-employer connection.
No explicit contract, minimum working hours vary non-systematically.
Their proportion has declined in the U.S. between 1970s and 1990s.

Possible **mental consequences**: 

 - job insecurity & unpredictability
 - low control over their work
 - perceiving themselves as disadvantaged
 
Most likely employers: firms in dynamic **competitive environments**.


## Independent contractors

Self-employed, no wage contract, responsible for their own tax arrangements.
U.S.: relatively little change since 1970s.


Their clients are not "vicariously liable" for contractors' actions (unlike vis-à-vis their employees) `r fa("arrow-alt-circle-right", fill = "#cc0033")` not required to provide fringe benefits and to pay unemployment compensation.

In the U.S., 38\% of employers **intentionally misclassify** their employees as "independent contractors" to **avoid payroll taxes**.


### Career

Many earn more than normal employees, but are less likely to receive health insurance and pensions.
"Portfolio workers" -- benefiting from moving between firms (e.g. Silicon Valley).
