---
title: "Labor market reforms in Europe"
description: |
  Lecture 4
#author: []
date: 2021-12-12
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    self_contained: false
    # code_folding: hide
creative_commons: CC BY
---


```{r setup, include=FALSE}
library(knitr)
library(here)
library(ymlthis)
library(tidyverse)
library(fontawesome)
library(readr)
library(ggthemes)
library(ggpubr)
library(RColorBrewer)
library(kableExtra)
library(data.table)
library(metathis)
library(xaringanthemer)
library(xaringanExtra)
library(data.table)
library(RefManageR)
library(plm)
library(mgcv)
library(mgcViz)
library(stargazer)
library(plotly)
library(htmlwidgets)
library(modelsummary)
```


```{r echo=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(collapse = TRUE,
                      fig.retina = 3)
```


::: {.define .article style="color: blue;"}
[Eichhorst, W., Marx, P., & Wehner, C. (2017). Labor market reforms in Europe: towards more flexicure labor markets?. *Journal for Labour Market Research, 51* (1), 1-17.](https://labourmarketresearch.springeropen.com/articles/10.1186/s12651-017-0231-7)
:::

## Inroduction

Asymmetrical impact of the 2008/2009 recession on young people and the subsequent wave of labor market reforms implemented particularly in countries hit by the crisis.

Europe: divide between permanent and temporary jobs.

Overview: 

 - showing trends in this form of labor segmentation;
 - overview of recent labor market reforms; 
 - further segmentation or mitigated divides in the labor market?
 - to what extent countries have embarked on the flexibility (or flexicurity) agenda and; has this led to a reduction of segmentation?

Goal of overcoming labor segmentation `r fa("arrow-alt-circle-right", fill = "#cc0033")` lowering employment protection.

**<span style="color:white;background-color:black">Flexicurity</span>** approach:

 - social protection and human capital investment compensating for low job security;
 - welfare state compensates for disadvantage in the labour market and facilitates mobility.

Initiatives to overcome labor segmentation: (so far) failed to produce noticeable changes.

Labor market segmentation: 

 - employment protection legislation (EPL); 
 - unemployment benefits;
 - active labor market policies (ALMP);
 - working-time flexibility

 
## Theoretical background

Labor market is only segmented if there are significant **mobility barriers**.

Temporary employment:

1. **Macro-economic efficiency**:

 - hampered productivity growth;
 - failed to increase employment levels (secure employment is substituted with temporary jobs)
 
2. **Workers' well-being**:

 - depressed health;
 
 - eligibility criteria for social insurance benefits often disadvantage workers with short and interrupted work records;
 
 - employment structures `r fa("arrows-alt-h", fill = "blue")` welfare state institutions `r fa("arrow-alt-circle-right", fill = "#cc0033")` labor market disadvantage `r fa("arrows-alt-h", fill = "blue")` opposed to the notion of flexicurity promoted by the European Commission
 
 
## Background: labor market segmentation in the EU

Focus: segmentation between workers with **permanent** and **temporary** employment contracts.

Part-time employment: 

 - mostly voluntary (except Greece, Italy, Spain & Slovakia);

 - Germany’s sizable apprenticeship system: 50\% of temporary contracts;
 
 - low shares (< 6\%): Bulgaria, Estonia, Latvia, Lithuania, Romania, the U.K.;
 
 - "55-64" group: less pronounced compared to the general workforce;
 
 - young workers: most problematic group
 
 
<span style="color:darkred">Mediterranean countries</span>: lack of labor demand `r fa("arrow-alt-circle-right", fill = "#cc0033")` insufficient number of secure positions for youth `r fa("arrow-alt-circle-right", fill = "#cc0033")` reducing temporary employment -- not a top priority.

**<span style="color:white;background-color:black">However</span>**, if temporary workers remain in insecure positions for a long time `r fa("arrow-alt-circle-right", fill = "#cc0033")` "trapped" in insecure and precarious jobs.

Temporary `r fa("arrow-alt-circle-right", fill = "#cc0033")` permanent after 10 years:

 - 100\% in Austria, Estonia and Germany; 
 
 - 50\% in Italy and Spain are still temps;
 
 - segmentation is more severe in member states with high **youth unemployment**.
 

## Measures against labor market segmentation

Strictly regulated labor markets: costly dismissals `r fa("arrow-alt-circle-right", fill = "#cc0033")` temporary contracts that allow for cheaper adjustment of the workforce.

1. **De-regulation of permanent contracts**: France, Greece, Italy, Portugal, Slovenia and Spain.

2. **Re-regulation of temporary contracts**: other countries.


### Dismissal protection reform

 - Greece, Portugal, Slovenia, Spain: reduced notice periods and severance pay for (newly hired) permanent workers

<span style="color:darkred">Italy</span>: employers do not have to reinstate workers dismissed for invalid economic reasons.

<span style="color:darkred">Spain</span>: less severance pay + more temporary workers' protection `r fa("arrow-alt-circle-right", fill = "#cc0033")` internal flexibility.

 - Finland, Poland, Sweden: high share of temps but no reforms
 
 
### Temporary employment

**Incentives**:

- making temporary contracts more expensive for employers; 
- offering monetary incentives to allow for transitions into permanent jobs 

<span style="color:darkred">Slovenia</span> 2013 labor market reform: 

 - raised employers’ unemployment insurance contributions for temporary jobs; 
 - if temporary job is converted into a permanent one `r fa("arrow-alt-circle-right", fill = "#cc0033")` employers are exempted from contributions for a limited time;
 - 2013 reform limited temporary contracts to 2 years
 
<span style="color:darkred">France</span>: employers are exempt from contributions for a limited period if they hire younger or older workers on **permanent contracts**.

<span style="color:darkred">Italy</span> higher contribution for temporary than permanent workers.

<span style="color:darkred">Spain</span> from 2015 onward: 

 - temporary workers receive a higher severance pay at the end of their contract;
 - payment corresponds to the salary of 12 days; much lower than for permanent workers
 
 
<span style="color:darkred">Netherlands</span> 2015: 3 `r fa("arrow-alt-circle-right", fill = "#cc0033")` 2 years


**Re-regulation** (opposite tendency):

<span style="color:darkred">Italy</span>: 

 - abolished the need to indicate a valid reason for temporary employment;
 - extended the maximum duration as well as the possible number of renewals (also in Spain and Greece);
 - **<span style="color:white;background-color:black">however</span>**, firms not allowed to use temporary contracts for **> 20\%** of the workforce and Spain
 
**Overall trends**: 

 - lower dismissal costs for permanent workers + moderate / ambiguous re-regulation of temporary contracts;
 
 - long-term declining trend in permanent hirings (Austria, Belgium and Luxembourg);
 
 - exceptions: Baltic states, Denmark, Romania
 
 - EPL for temps: significantly related to a decrease in temporary employment (when not controlling for country- and year-effects)


### Unemployment benefits

**Flexicurity** rationale: 

 - access to social benefits
 
 - avoiding double disadvantage: losing job + lacking access to unemployment insurance benefits
 
 
<span style="color:darkred">Italy</span>:

 - 2013: **Assicurazione Sociale per l’Impiego* (ASPI) `r fa("arrow-alt-circle-right", fill = "#cc0033")` more generous and encompassing system;
 
 - 2015: (NASPI) more permissive eligibility criteria (13 weeks of contributions in the 4 years before unemployment);
 
 - new unemployment benefit scheme for atypical employees
 
 
<span style="color:darkred">France</span>:
4 months of employment over the last 28 months `r fa("arrow-alt-circle-right", fill = "#cc0033")` eligible for 4-months+ unemployment insurance benefits.


<span style="color:darkred">Spain</span>:

 - more access to unemployment benefits for temporary and younger workers as well as self-employed;
 
 - stronger work incentives and activation policies for the unemployed

2007-2015 **Overall trend**: declining share of short-term unemployed receiving benefits 


### Active labor market policies (ALMPs)

**Segmented labor markets**: vulnerable workers have a higher risk of becoming unemployed `r fa("arrow-alt-circle-right", fill = "#cc0033")` entering into contact with public employment services (PES).

Effective ALMPs and PESs `r fa("arrow-alt-circle-right", fill = "#cc0033")`:

 - quick and sustainable re-employment;
 - subsidizing hirings on a temporary basis (has its downsides as well); 
 - incentives for conversions into permanent positions
 
Increased **training expenditure**: Austria, Finland, Sweden

Increased **expenditure per unemployed**: Germany (note: declining number of unemployed), Denmark, Sweden, Estonia

**Low-skilled workers**: 

 - prime beneficiaries of training;
 - but, *minority* of all temporary workers in all countries
 
> **1/4** ~ **1/3** of all temporary workers have tertiary education `r fa("arrow-alt-circle-right", fill = "#cc0033")` unlikely to benefit from ALMP intervention (e.g. training) `r fa("arrow-alt-circle-right", fill = "#cc0033")` still insufficient jobs for skilled workers

## Conclusions

EU-states:

 - policy initiatives `r fa("arrow-alt-circle-right", fill = "#cc0033")` intending to overcome labor market divides;
 
 - **<span style="color:white;background-color:black">however</span>**, temporary employment `r fa("arrow-alt-circle-up", fill = "#cc0033")`; youth unemployment remains high
 
1. Temporary employment can serve as **bridge** to permanent jobs `r fa("arrow-alt-circle-right", fill = "#cc0033")` prohibitive regulation -- not an option

2. **Deregulation** of permanent contracts `r fa("arrow-alt-circle-up", fill = "#cc0033")` not automatically translating into encompassing job security

<span style="color:darkred">France</span> **insecurity bonuses**: employers paying a bonus to the worker (10\% of the total gross wage) if the employment relationship is not continued after the contract has expired.

Main **challenges**:

 - political preferences: reducing social rights of "insiders" `r fa("arrow-alt-circle-right", fill = "#cc0033")` difficult to organize democratic majorities; 
 
 - expanding social protection for vulnerable workers `r fa("arrow-alt-circle-right", fill = "#cc0033")` resistance from the majority with a low unemployment risk (net payers);
 
 - insufficient labor demand `r fa("arrow-alt-circle-right", fill = "#cc0033")` workers and policy-makers view temporary employment as the **lesser of two evils**
