---
title: "Labor market reforms in Europe"
description: |
  Lecture 4
#author: []
date: 2021-12-12
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    self_contained: false
    # code_folding: hide
creative_commons: CC BY
---


```{r setup, include=FALSE}
library(knitr)
library(here)
library(ymlthis)
library(tidyverse)
library(fontawesome)
library(readr)
library(ggthemes)
library(ggpubr)
library(RColorBrewer)
library(kableExtra)
library(data.table)
library(metathis)
library(xaringanthemer)
library(xaringanExtra)
library(data.table)
library(RefManageR)
library(plm)
library(mgcv)
library(mgcViz)
library(stargazer)
library(plotly)
library(htmlwidgets)
library(modelsummary)
```


```{r echo=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(collapse = TRUE,
                      fig.retina = 3)
```


::: {.define .article style="color: blue;"}
[Eichhorst, W., Marx, P., & Wehner, C. (2017). Labor market reforms in Europe: towards more flexicure labor markets?. *Journal for Labour Market Research, 51* (1), 1-17.](https://labourmarketresearch.springeropen.com/articles/10.1186/s12651-017-0231-7)
:::

## Inroduction

Asymmetrical impact of the 2008/2009 recession on young people and the subsequent wave of labor market reforms implemented particularly in countries hit by the crisis.

Europe: divide between permanent and temporary jobs.

Overview: 

 - showing trends in this form of labor segmentation;
 - overview of recent labor market reforms; 
 - further segmentation or mitigated divides in the labor market?
 - to what extent countries have embarked on the flexibility (or flexicurity) agenda and; has this led to a reduction of segmentation?

Goal of overcoming labor segmentation `r fa("arrow-alt-circle-right", fill = "#cc0033")` lowering employment protection.

**<span style="color:white;background-color:black">Flexicurity</span>** approach:

 - social protection and human capital investment compensating for low job security;
 - welfare state compensates for disadvantage in the labour market and facilitates mobility.

Initiatives to overcome labor segmentation: (so far) failed to produce noticeable changes.

Labor market segmentation: 

 - employment protection legislation (EPL); 
 - unemployment benefits;
 - active labor market policies (ALMP);
 - working-time flexibility

 
## Theoretical background

Labor market is only segmented if there are significant **mobility barriers**.

Temporary employment:

1. **Macro-economic efficiency**:

 - hampered productivity growth;
 - failed to increase employment levels (secure employment is substituted with temporary jobs)
 
2. **Workers' well-being**:

 - depressed health;
 
 - eligibility criteria for social insurance benefits often disadvantage workers with short and interrupted work records;
 
 - employment structures `r fa("arrows-alt-h", fill = "blue")` welfare state institutions `r fa("arrow-alt-circle-right", fill = "#cc0033")` labor market disadvantage `r fa("arrows-alt-h", fill = "blue")` opposed to the notion of flexicurity promoted by the European Commission
 
 
## Background: labor market segmentation in the EU

Focus: segmentation between workers with **permanent** and **temporary** employment contracts.

Part-time employment: 

 - mostly voluntary (except Greece, Italy, Spain & Slovakia);

 - Germany’s sizable apprenticeship system: 50\% of temporary contracts;
 
 - low shares (< 6\%): Bulgaria, Estonia, Latvia, Lithuania, Romania, the U.K.;
 
 - "55-64" group: less pronounced compared to the general workforce;
 
 - young workers: most problematic group
 
 
<span style="color:darkred">Mediterranean countries</span>: lack of labor demand `r fa("arrow-alt-circle-right", fill = "#cc0033")` insufficient number of secure positions for youth `r fa("arrow-alt-circle-right", fill = "#cc0033")` reducing temporary employment -- not a top priority.

**<span style="color:white;background-color:black">However</span>**, if temporary workers remain in insecure positions for a long time `r fa("arrow-alt-circle-right", fill = "#cc0033")` "trapped" in insecure and precarious jobs.

Temporary `r fa("arrow-alt-circle-right", fill = "#cc0033")` permanent after 10 years:

 - 100\% in Austria, Estonia and Germany; 
 
 - 50\% in Italy and Spain are still temps;
 
 - segmentation is more severe in member states with high **youth unemployment**.
 

## Measures against labor market segmentation

Strictly regulated labor markets: costly dismissals `r fa("arrow-alt-circle-right", fill = "#cc0033")` temporary contracts that allow for cheaper adjustment of the workforce.

1. **De-regulation of permanent contracts**: France, Greece, Italy, Portugal, Slovenia and Spain.

2. **Re-regulation of temporary contracts**: other countries.


### Dismissal protection reform

 - Greece, Portugal, Slovenia, Spain: reduced notice periods and severance pay for (newly hired) permanent workers

<span style="color:darkred">Italy</span>: employers do not have to reinstate workers dismissed for invalid economic reasons.

<span style="color:darkred">Spain</span>: less severance pay + more temporary workers' protection `r fa("arrow-alt-circle-right", fill = "#cc0033")` internal flexibility.

 - Finland, Poland, Sweden: high share of temps but no reforms
 
 
### Temporary employment

Incentives:

- making temporary contracts more expensive for employers; 
- offering monetary incentives to allow for transitions into permanent jobs 

<span style="color:darkred">Slovenia</span> 2013 labor market reform: 

 - raised employers’ unemployment insurance contributions for temporary jobs; 
 - if temporary job is converted into a permanent one `r fa("arrow-alt-circle-right", fill = "#cc0033")` employers are exempted from contributions for a limited time;
 - 2013 reform limited temporary contracts to 2 years
 
<span style="color:darkred">France</span>: employers are exempt from contributions for a limited period if they hire younger or older workers on **permanent contracts**.

<span style="color:darkred">Italy</span> higher contribution for temporary than permanent workers.

<span style="color:darkred">Spain</span> from 2015 onward: 

 - temporary workers receive a higher severance pay at the end of their contract;
 - payment corresponds to the salary of 12 days; much lower than for permanent workers
 
 
<span style="color:darkred">Netherlands</span> 2015: 3 `r fa("arrow-alt-circle-right", fill = "#cc0033")` 2 years
