---
title: "Coordination as a political problem"
description: |
  Lecture 2
# author:
#   - name: Kostiantyn Ovsiannikov
date: 2021-11-09
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    self_contained: false
    # code_folding: hide
creative_commons: CC BY
---


```{r setup, include=FALSE}
library(knitr)
#library(summarytools)
library(here)
library(ymlthis)
library(tidyverse)
library(fontawesome)
library(readr)
library(ggthemes)
library(ggpubr)
library(RColorBrewer)
library(kableExtra)
library(data.table)
library(metathis)
library(xaringanthemer)
library(xaringanExtra)
library(data.table)
library(RefManageR)
library(plm)
library(mgcv)
library(mgcViz)
library(stargazer)
library(plotly)
library(htmlwidgets)
library(modelsummary)
```


```{r echo=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(collapse = TRUE,
                      fig.retina = 3)
```


```{r, load_refs, include=FALSE, cache=FALSE}
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "alphabetic",
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib(here("_posts", "lecture-1", "Labor-Dualism.bib"), check = FALSE)
```


<!-- ## Article discussed -->

::: {.define .article style="color: blue;"}
[Thelen, K., & Kume, I. (2006). Coordination as a political problem in coordinated market economies. *Governance, 19* (1), 11-42.](https://onlinelibrary.wiley.com/doi/abs/10.1111/j.1468-0491.2006.00302.x)
:::

## Gist

**Purpose**: to examine the differences in how *employer coordination* has been organized in Sweden, Germany, and Japan in the area of industrial relations.

**Argument**: intensification of *cooperation between labor and management* in some firms and industries has paradoxically had <span style="color:darkred">deeply destabilizing collateral effects</span> that have undermined these systems as they were traditionally constituted.


## Background

Recent **Varieties of Capitalism** (VoC): industrial relations depend not only upon labor organization and strength, but also on the *coordination between employers*.

Coordinated market economies (CMEs):

 - Continental Europe
 - Japan

Liberal market economies (LMEs):

 - U.S.
 - U.K.

Some argue that globalization and deregulation will lead to convergence – all capitalist systems will become liberal.

**<span style="color:white;background-color:black">VoC opinion</span>**:

 - **continued diversity** and even stronger divergence between LMEs and CMEs
 - employers in the developed democracies will build on and deepen previous sources of **comparative institutional advantage**
 
    - employers in the <span style="color:darkred">U.S.</span> and the <span style="color:darkred">U.K.</span> -- will seek **deregulation**;
    - employers in <span style="color:darkred">Germany</span>, <span style="color:darkred">Sweden</span>, and <span style="color:darkred">Japan</span> -- will hold to high degree of **labor cooperation**, strong investment in workers' **skill acquisition**, and peaceful plant relations.
 
**Basic logic**: employers have organized their competitive strategies around these institutions `r fa("arrow-alt-circle-right", fill = "#cc0033")` will not abandon them in the face of new market pressures.


### Challenges to the VoC framework

> Do institutions really resist deregulation thanks to employers’ genuine interest?

Alternative explanations: 

 - institutional inertia 
 - successful union strategies
 
Significant tensions in <span style="color:darkred">Sweden</span>, <span style="color:darkred">Germany</span> (declining union membership) and <span style="color:darkred;class="bg-danger>Japan</span> (departure from seniority wages and lifetime employment). 
Thus, not only “coordination”, but also **struggles** matter.


Contemporary market conditions:

 - not only new conflicts between labor and capital (as is widely known and theorized) 
 - new tensions **among firms and industries** that are differently situated in domestic and international markets
 - employers' coordination as a **political problem**
 
> Intensification of cooperation between labor and management in some firms and industries (that the VoC literature correctly emphasizes) has paradoxically deep **destabilizing** collateral effects.


## Germany

Background: organizational disarray in the Association of German Metalworking firms (*Gesamtmetall*) and its union counterpart (*IG Metall*), which together have played the key flagship role in Germany’s *de facto* system of **pattern bargaining** over most of the postwar period.

Firms calling for greater flexibility to adapt central bargains to the needs of individual firms.

1960s: **collective bargaining** was covering 80\% of all firms accounting for 90\% of all workers.

**Union organization** in <span style="color:darkred">Germany</span>: around 30\% of all workers throughout the postwar period vs. <span style="color:darkred">Swedish</span> unions’ 80\%.

**<span style="color:white;background-color:black">However</span>**, German employers have traditionally been extremely well-organized `r fa("arrow-alt-circle-right", fill = "#cc0033")` contracts signed by the employer associations cover all workers in the member firms, regardless of their status as union members or not `r fa("arrow-alt-circle-right", fill = "#cc0033")` stabilized encompassing bargaining

**Large export firms** `r fa("arrows-alt-h", fill = "blue")` sizable sector of small and medium-sized firms (*Mittelstand*):

 - large firms dominated the employer association, but also bore the **burden of industrial conflicts** `r fa("arrow-alt-circle-right", fill = "#cc0033")` settlements with the unions on a range of issues such as wages

Some of the most <span style="color:darkred">destabilizing trends</span>: intensified **cooperation between labor and capital** at the plant level (especially in larger firms).

Clear tensions since 1980s: negotiations over **working hours' reduction** (down to 35):

 - larger firms -- allowed greater flexibility in working times, expanding shift work or devising differentiated working time arrangements `r fa("arrow-alt-circle-right", fill = "#cc0033")` **offsetting the costs** of such reductions
 - smaller companies -- not able to make same kinds of adjustments `r fa("arrow-alt-circle-right", fill = "#cc0033")` were **hit more heavily** by the sheer costs of reduced working hours
 
Subsequent collective bargaining rounds in the 1990s `r fa("arrow-alt-circle-right", fill = "#cc0033")` exacerbated **problems of solidarity** on the employer side.

German producers: facing much **stiffer competition** from high-quality (lower-cost) producers in Japan and elsewhere.
Late 1990s: bargaining rounds that featured remarkable displays of **disunity on the part of metalworking employers** (that became risk-averse).

Weaker firms: <span style="color:darkred">opting out</span> of the industry-level contract altogether `r fa("arrow-alt-circle-right", fill = "#cc0033")` shrinking number of firms participating in central bargaining.

Union leaders' ability to conclude binding and encompassing agreements `r fa("arrows-alt-h", fill = "blue")` continued organizational viability of employers.
Unions' achievements were muted, frequently accompanied by expressions of concern about the **weakness of employers' association**.

**Lack of unity and discord**: prominent features of the IG Metall’s internal politics at present.

In 2003, the *IG Metall* strike (over reducing the working time in East Germany from 38 to 35 hours, as in the West) failed for the first time since 1950s.

Employers such as *Gesamtmetall*’s head Kannegiesser explicitly defend the **collective bargaining** system for the “order” and “predictability” that it helps to impose on the labor market.


## Japan

Postwar system of **"lifetime employment"** (sh&#363;shin koy&#333;) + **seniority wages** (nenk&#333; joretsu): incentive mechanisms for workers.

Japanese management practices operate on a **firm level** (rather than on industrial, as in Germany).

Importance of <span style="color:darkred">coordination</span> (rather than competition) between employers.

Lifetime employment only works if workers start at low wages, and then gradually acquire skills that make them increasingly productive and valuable to the company.

Management-labor coordination is less formal than in Germany.
Conducted via the seasonal *Shunt&#333;* **wage bargaining** system.

Since 1960s: the metal industry had become the main pattern-setting industry.  Unions, organized in the IMF-JC (International Metal Workers’ Federation Japan Council), have traditionally demanded “economically rational” wage increases.

Recent years: intensified questioning of the traditional wage system. 
Employer organizations such as *Keidanren* (Federation of Economic Organizations) and **Nikkeiren** have criticized seniority wages as rigid and harmful for Japanese firms. 
Reasons:

1. **Demographic change** (aging workforce) has increased the number of older workers, and with that, increased also the total wage costs that firms must bear.

2. **Seniority wage system change**: came closer to an automatic wage raise for many workers despite different job capabilities among workers of the same age.

3. **“IT revolution”**: changed the kinds of skills that firms need. 
Traditional on-the-job training over a worker’s career may be less attractive than hiring engineers with the relevant skills from outside the company.

Thus, many companies have moved to abolish the seniority wage system, often replacing this with **performance-based wages** (*seikashugi*).

However, as in Germany, one finds in Japan evidence of continued employer support for both lifetime employment practices and substantial aspects of traditional wage systems. 
1995 *Nikkeiren* report <span style="color:darkred">“A Japanese Management System in the New Era”</span>.
Urging firms to distinguish among three different types of workers: 

- long-term employed workers with accumulated skills; 

- workers with professional skills; 

- short-term/flexibly employed workers.


<span style="color:darkred">Toyota</span> management: 

 - if Toyota (profiting from the weak yen) raised the regular wage `r fa("arrow-alt-circle-right", fill = "#cc0033")` spilling over into other industries `r fa("arrow-alt-circle-right", fill = "#cc0033")` exacerbating the high cost problem;
 
 - alternative suggestion: redistributing increased company profit in the form of **bonus payments**, which are less linked to *Shunt&#333;*;
 
 - **result**: Toyota increased its regular wage by a larger margin than other companies in the metal sector `r fa("arrow-alt-circle-right", fill = "#cc0033")` abdicating its role as benchmark and pattern setter;
 
 - Toyota is convinced that **secure employment** should be the basis of its competitiveness.

<span style="color:darkred">Toyota</span> chairman Hiroshi Okuda in 1999: 

  > “Even if there are redundant workers, it is management’s responsibility to seek new business opportunities to redeploy them.
Employers without the entrepreneurship to find such business opportunities should retire.” 

The chairman of the board of <span style="color:darkred">Fujitsu</span> concurred with Okuda, citing his company’s successful efforts to retain workers through **retraining** and **redeployment**.
Fujitsu management: it is important to evaluate workers’ **potential ability** and “work process” independently of “outcome”.

**<span style="color:white;background-color:black">On contrary</span>**, pioneer of the performance-based system, <span style="color:darkred">Takeda Chemical</span> (the largest pharmaceutical company in Japan), found it difficult to keep employees committed to the company.

1990s: <span style="color:darkred">Takeda Chemical</span> and <span style="color:darkred">Mitsui Mining and Smelting</span> adopted performance-based wages, leaving the *Shunt&#333;* framework altogether.

**The major problem**: how to make individualistic wages compatible with the **group-oriented** work organization of Japanese firms.

<span style="color:darkred">Nissan</span> case:

 - taken over by the French automaker <span style="color:darkred">Renault</span> in 1998 due to its financial difficulties;
 
 - went through a very tough restructuring with massive dismissals (in the form of voluntary retirements);
 
 - however, after achieving a V-curve recovery, president Carlos Ghosn maintained a bottom-up decision-making style, (which is regarded as very Japanese) and made it clear that the company’s main strength should reside in **worker commitment** to the firm; 
 
 - wage bargaining for 2001, and in the context of concerns about revitalizing work organization: Nissan unilaterally declared that it would pay the **full amount of bonuses** demanded by the union.


**<span style="color:white;background-color:black">However</span>**, not all companies can accomplish this same balance (traditional employment practices + fine-tuning wage systems). 

**New divide** among core firms in the Japanese economy:

 - stronger firms (such as Toyota); 
 
 - weaker firms: traditional practices (including laying off workers) `r fa("arrow-alt-circle-right", fill = "#cc0033")` straightforward efforts to reduce labor costs;
 
 - for example, <span style="color:darkred">Mazda</span>: seniority-based wages `r fa("arrow-alt-circle-right", fill = "#cc0033")` performance-based system.
 Reason: it could no longer afford Toyota-style wage and employment practices. 
 Radical reform of the wage system `r fa("arrow-alt-circle-right", fill = "#cc0033")` **demoralizing** workers, not improving company competitiveness.
 
 
Increasing adoption of performance-based wage systems `r fa("arrow-alt-circle-right", fill = "#cc0033")` difficult to maintain the traditional *Shunt&#333;* system:

  - *Shunt&#333;* coordination `r fa("arrows-alt-h", fill = "blue")` leading firms acting as pattern setters both within their own industry and also with an eye to other industries;
  - **performance-based wages** interfere with this. 
  Difficulty of **coordinating wages** `r fa("arrows-alt-h", fill = "blue")` company profits begin to play a prominent role in wage determination.
  

## Sweden

### Background

High level of **centralization**.

1956-1983: **broad wage settlements** blue-collar union confederation *Landorganisationen* (LO) `r fa("arrows-alt-h", fill = "blue")` *National Confederation of Swedish Employers* (SAF).

1980-90s: **collapse of the system**, as SAF eliminated its own bargaining unit.

**<span style="color:white;background-color:black">However</span>**, from 1990s on: **re-equilibration** of wage bargaining instead of decentralization.

Currently: highly **coordinated bargaining** across industries within the export sector.
Much stronger **cooperation** than ever before between blue-collar unions and white-collar workers. 

**<span style="color:white;background-color:black">However</span>**: much looser coordination between the exposed and sheltered sectors of the economy.

`r fa("arrow-alt-circle-right", fill = "#cc0033")` growing **wage differentiation** within and especially between these broad bargaining clusters `r fa("arrow-alt-circle-right", fill = "#cc0033")` significant compression of wages across the economy.


### The case

Swedish employers: originally agreed to centralized bargaining `r fa("arrow-alt-circle-right", fill = "#cc0033")` ensuring wage moderation. 
However, over time wage settlements became **rigid** and highly **inflationary**.

Before 1960s: centralized bargaining `r fa("arrows-alt-h", fill = "blue")` wage solidarism `r fa("arrow-alt-circle-right", fill = "#cc0033")` equalizing pay across industries and plants (**"equal pay for equal work"** regardless of a firm’s profitability) `r fa("arrow-alt-circle-right", fill = "#cc0033")` shift in resources (including labor) out of declining and into more competitive industries.

From late 1960s: wage solidarity took an **egalitarian** turn `r fa("arrow-alt-circle-right", fill = "#cc0033")` compressing relative wages across the board. 

`r fa("arrow-alt-circle-right", fill = "#cc0033")` compensating low-paid groups (public sector and low-skilled industrial workers) for wage drift enjoyed by skilled industrial workers `r fa("arrow-alt-circle-right", fill = "#cc0033")` chronic, institutionalized **ratcheting up** of wages.
  
1980s and 1990s: negative effects felt especially in the **engineering industry** `r fa("arrow-alt-circle-right", fill = "#cc0033")` employer association in that sector (*Verkstadsf&#246;reningen*, or VF; later *Verstadsindustrier*, or VI) led the drive for **bargaining decentralization**.
  
VF/VI opposed the new wage-leveling clauses `r fa("arrow-alt-circle-right", fill = "#cc0033")` **no resonance** among employers in other industries that were less affected `r fa("arrow-alt-circle-right", fill = "#cc0033")` 1983: VI withdrew from traditional bargaining altogether `r fa("arrows-alt-h", fill = "blue")` separate deal with the metalworkers’ union, Metall `r fa("arrow-alt-circle-right", fill = "#cc0033")` **decisive blow** to the traditional nationally-coordinated system.

Employers in the **engineering industry** led the charge against the old system. 
Rather than "neoliberal employer offensive" against the traditional Swedish system,
**cross-class realignment** `r fa("arrow-alt-circle-right", fill = "#cc0033")` bringing together employers and unions in the export sectors (especially in the engineering industry), at the expense of traditional, more encompassing forms of coordination and solidarity.

1983: initial break with centralized bargaining -- not through conflict, but through a **deal**. 
VF offered the metalworkers’ union (Metall) a **wage increase** above what the union had demanded `r fa("arrows-alt-h", fill = "blue")` **decoupling** negotiations from the peak bargain and eliminating contractual provisions that compressed the wages of skilled and unskilled workers.

**Engineering workers** -- thinking of public-sector workers as **"pay parasites"** `r fa("arrow-alt-circle-right", fill = "#cc0033")` resented their ability to **free-ride** on the productivity gains in engineering through the compensation clauses.

Furthermore, interoccupational **leveling clauses** (between skilled and unskilled workers) -- a problem for *Metall*.
Technological changes in engineering `r fa("arrow-alt-circle-right", fill = "#cc0033")` blurring the line between **skilled blue-collar** and **white-collar** occupations `r fa("arrow-alt-circle-right", fill = "#cc0033")` high-skill metalworkers working in jobs similar to that of members of the separate white-collar union, *Svenska Industritj&#228;nstemannaf&#246;rbundet* (SIF). 
**<span style="color:white;background-color:black">But</span>**, wage compression within *Metall* `r fa("arrow-alt-circle-right", fill = "#cc0033")` receiving much lower wages than white-collar workers.

VI’s break with the system `r fa("arrow-alt-circle-right", fill = "#cc0033")` **uncertainty** `r fa("arrows-alt-h", fill = "blue")` demands from firms like <span style="color:darkred">Volvo</span>, <span style="color:darkred">Ericsson</span>, and <span style="color:darkred">ABB</span> for further **decentralization** to the plant level. 

Plus, *SAF* dismantling of its own bargaining apparatus in 1990.
Employers continued to oppose any recentralization of national bargaining under the auspices of the *LO* and *SAF*.

But, other firms and sectors advocating for more **intermediate solution** of industry-level negotiations `r fa("arrow-alt-circle-right", fill = "#cc0033")` both full decentralization to the plant level and return to national-level bargaining are **impossible**.

<span style="color:darkred">1995 bargaining round</span>: 

 - highlighting potential pitfalls (for employers) of a **lack of cooperation** across industries within private sector manufacturing;
 
 - underscoring disadvantages of completely uncoordinated (industry-level) bargaining in a context in which unions are capable of backing up their demands with actions that export-dependent industries find extremely expensive in highly competitive and tightly integrated global markets;
 
 - reconfiguration of bargaining that is based on: 
 
    * cooperation across industries;
    
    * cooperation between labor and capital within the export sector; 
    
    * abandoning encompassing forms of coordination and solidarity at the national level 


1996: *Metall* joined with its white-collar counterpart (*SIF*) and other unions in export sectors `r fa("arrow-alt-circle-right", fill = "#cc0033")` invited other employers to engage in joint negotiations over wage formation.

1997: 8 unions and 12 employer associations `r fa("arrow-alt-circle-right", fill = "#cc0033")` new "Agreement on Industrial Development and Wage Formation":

 - establishing **mediators** to promote peaceful compromise within industries, as well as to broker coordination across industries in the exposed sector as a whole;
 
 - importance for **unions**, but especially for *Metall*: prompting the *VI* to abandon its earlier efforts to push for full decentralization, ending years of struggle in this key industry over the structure of negotiations;
 
 - importance for **employers**: the new conflict mediation procedures that lie at the center of the agreement, crucial to managing their heightened vulnerability to interruptions in production, including strikes; 

 - **decoupling** bargaining in the exposed sector from that in the sheltered sector, facilitating a closer interface between white- and blue-collar contracts in industry;
 
 - completely sidelining the national trade union confederation *LO* `r fa("arrow-alt-circle-right", fill = "#cc0033")` *LO* unenthusiastic about the industry agreement;
 
 - the **export sector** (on both the labor and employer sides) opposed any government machinery (be it *SAF* or *LO*) that would circumvent or supersede the successful procedures they had worked out for themselves, a position that was respected in the legislation that ultimately emerged in 2000;
 
 - public-sector workers concerned with preserving the right to strike `r fa("arrow-alt-circle-right", fill = "#cc0033")` hurried to conclude a voluntary agreement similar to the industry agreement `r fa("arrow-alt-circle-right", fill = "#cc0033")` keeping themselves out of government mediation machinery;

 - prospect of state regulation `r fa("arrow-alt-circle-right", fill = "#cc0033")` greater attempts at self-regulation `r fa("arrow-alt-circle-right", fill = "#cc0033")` end of 2001: 16 agreements covering over half of the entire workforce
 
 
2000: a new national Mediation Authority (*Medlingsinstitutet*) was established,
replacing the old, weaker National Conciliator’s Office:

 - intended only to **complement** the voluntary industry agreements;
 
  - promotes and works around existing agreements and covers those industries (**transportation** and **retail** trade among them) that have no such voluntary agreement;
  
  - mediators only have power of **persuasion**, and cannot impose a solution. 
  Less powerful than its **Danish** counterpart, despite similar goals -- to avoid industrial strife and especially to facilitate coordination across industries behind the lead of the exposed sectors
  

### Summary

**Employer coordination** has proved more resilient than adherents to a strict neoliberal offensive thesis might have predicted. 

Full decentralization to the plant level (as advocated in the 1980s and 1990s by prominent firms such as ABB, Volvo, and Ericsson) has been averted, and there is significant coordination across the industry sector as a whole.

Confederal level (with white- and blue-collar workers negotiating separately) `r fa("arrow-alt-circle-right", fill = "#cc0033")` coordination across the blue-/ white-collar divide and across export industries.

Since 1990s: union alliances based upon class increasingly replaced by **constellations based upon sector**.

**Renegotiation** of the terms of coordination:

1. Confederal bargaining `r fa("arrow-alt-circle-right", fill = "#cc0033")` industry-level bargaining: employers sloughed off several layers of coordination dragging on their ability to **restructure production** and to maneuver effectively in increasingly competitive international markets. 

*VI*’s initial withdrawal from peak negotiations `r fa("arrow-alt-circle-right", fill = "#cc0033")` blow to solidaristic wage bargaining as traditionally practiced in Sweden. 

Not that much neoliberal offensive by employers against labor, but the **intensified cooperation between labor and capital** in the metalworking industry at the expense of solidarity and coordination across sectors.

2. **Relinking** metal industry to other export industries, thus bringing back coordination organized around what the industry sector as a whole can bear. 

**Significant shift**:

 - *LO* relegated to the distant sidelines when it comes to wage negotiations; 
 
 - wage solidarity on the old terms is irretrievably gone;
 
 - employers are "dead against" any return to national-level bargaining; 
  
 - employers have been massively aided by the most **powerful industrial unions** (especially *Metall*), which have their own reasons to support the arrangements that rest on a new form of accommodation between labor and capital within **key export
industries**

## Conclusion

Recent market developments `r fa("arrow-alt-circle-right", fill = "#cc0033")` intensification of **cooperation** between labor and management `r fa("arrow-alt-circle-right", fill = "#cc0033")` weakening traditional forms of coordination among employers, undermining solidarity among workers and the system of **institutional coordination**.

Common trends: 

 - new, more virulent forms of **intensified cooperation** between management and labor; 
 
 - **<span style="color:white;background-color:black">but</span>**, growing gap between practices and outcomes within the new core and those outside it
