---
title: "Lecture 2"
description: |
  Defining non-standard employment
# author:
#   - name: Kostiantyn Ovsiannikov
date: 2021-11-01
output:
  distill::distill_article:
    self_contained: false
creative_commons: CC BY
---


```{r setup, include=FALSE}
library(knitr)
library(summarytools)
library(here)
library(ymlthis)
library(tidyverse)
library(fontawesome)
library(readr)
library(ggthemes)
library(ggpubr)
library(RColorBrewer)
library(kableExtra)
library(data.table)
library(metathis)
library(xaringanthemer)
library(xaringanExtra)
library(data.table)
library(RefManageR)
library(plm)
library(mgcv)
library(mgcViz)
library(stargazer)
library(plotly)
library(htmlwidgets)
library(modelsummary)
```


```{r echo=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(collapse = TRUE,
                      fig.retina = 3)
```


```{r, load_refs, include=FALSE, cache=FALSE}

BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "alphabetic",
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib(here("_posts", "lecture-2", "Labor-Dualism.bib"), check = FALSE)
```

```{r echo=FALSE, include=FALSE}
capture.output(print(myBib["kalleberg2000ars"]))
```


::: {.define .article style="color: blue;"}
Kalleberg, A. L. (2000). Nonstandard employment relations: Part-time, temporary and contract work. *Annual review of sociology, 26*(1), 341-365.
:::


# Background

1970s' **oil crises** `r fa("arrow-alt-circle-right", fill = "#cc0033")` sluggish economic growth `r fa("arrow-alt-circle-right", fill = "#cc0033")` high unemployment in the West `r fa("arrow-alt-circle-right", fill = "#cc0033")` inability to generate full employment for everyone `r Citep(myBib, "emmenegger2009joesp", .opts = list(cite.style = "authoryear", style = "html"))`.

Nonstandard employment relations are not new.
In fact, internal labor markets in the post-World War II period were more of an **exception** than a regularity.



