---
title: "Institutionalizing Dualism"
description: |
  Lecture 3
# author:
#   - name: Kostiantyn Ovsiannikov
date: 2021-11-27
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    self_contained: false
    # code_folding: hide
creative_commons: CC BY
---


```{r setup, include=FALSE}
library(knitr)
library(here)
library(ymlthis)
library(tidyverse)
library(fontawesome)
library(readr)
library(ggthemes)
library(ggpubr)
library(RColorBrewer)
library(kableExtra)
library(data.table)
library(metathis)
library(xaringanthemer)
library(xaringanExtra)
library(data.table)
library(RefManageR)
library(plm)
library(mgcv)
library(mgcViz)
library(stargazer)
library(plotly)
library(htmlwidgets)
library(modelsummary)
```


```{r echo=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(collapse = TRUE,
                      fig.retina = 3)
```


::: {.define .article style="color: blue;"}
[Palier, B., & Thelen, K. (2010). Institutionalizing dualism: Complementarities and change in France and Germany. *Politics & Society, 38* (1), 119-148.](https://journals.sagepub.com/doi/abs/10.1177/0032329209357888)
:::

## Argument

> Contrary to theoretical accounts that suggest that institutional complementarities support stability and institutional reproduction, the authors argue that the linkages across these realms have helped to translate employer strategies that originated in the realm of industrial relations into a stable, new, and less egalitarian model with state support.


Centralized collective bargaining `r fa("arrow-alt-circle-right", fill = "#cc0033")` firm-based bargaining.

**Labor markets**: not solely deregulated, but the number of "atypical" or "nonstandard" employment relationships + working poor -- risen sharply in recent years.

Significant changes `r fa("arrows-alt-h", fill = "blue")` high degree of formal institutional stability

- negotiated and "sold" politically as ways of **preserving**, not undermining, traditional arrangements and the social order they reflect and represent.


**Focus of the article**: 3 institutional arenas (industrial relations, labor markets, and social protection) in the political economies of <span style="color:darkred">Germany</span> and <span style="color:darkred">France</span>.

Some authors: see it as liberalization and a move toward the American model.

This article' **argument**: politics and the outcomes `r fa("arrow-alt-circle-right", fill = "#cc0033")` institutionalization of **<span style="color:white;background-color:black">new forms of dualism</span>** (less harsh than the American model but also distinctly less egalitarian than before). 

**Cross-class coalitions** `r fa("arrow-alt-circle-right", fill = "#cc0033")` European countries not succumbing to liberalization, but also helping to promote **dualization**. 

France \& Germany: 

 - actors in the "core" economy well-positioned to defend traditional institutions and practices for themselves, but no longer serving as leaders;
 
 - able to resist outright liberalization, but especially vulnerable to dualization


More recent work on "dualization": **macro-political** perspective.   
The theory by <span style="color:blue">David Rueda</span>: 

 - labor market **insiders** and **outsiders** have distinct (and in many ways contradictory) preferences with respect to labor market policy;
 
 - insiders: preferring strong employment protections over more flexibility to ease the reentry of outsiders into the labor market; 
 
 - outsiders: preferring the opposite;

 - **social-democratic** governments `r fa("arrow-alt-circle-right", fill = "#cc0033")` interests of **insiders** over those of outsiders;
 
 - corporatist arrangements: do not stimulate social democrats to bring the interests of outsiders on board;
 
 - social democracy: not a problem (as Rueda argues) but a part of the solution 
(to resist dualism);

 - dualist trends: far more pronounced in the continental European countries than in the Nordic countries, where social democratic parties are stronger and in some cases near hegemonic;
 
 - **political coalitions** `r fa("arrow-alt-circle-right", fill = "#cc0033")` labor market policy, industrial relations, social protection;
 
 - institutionalization of dualism `r fa("arrows-alt-h", fill = "blue")` changes in industrial relations, labor market policy, and welfare state reforms. 
 
Theoretical lenses: political economy `r fa("arrows-alt-h", fill = "blue")` **institutional complementarities**. 

Core industrial sectors `r fa("arrow-alt-circle-right", fill = "#cc0033")` "standard employment relationships":

 - <span style="color:darkred">Germany</span>: large export firms in manufacturing (e.g., automobiles and machine
tools);

 - <span style="color:darkred">France</span>: nationalized firms like *Renault* + large public enterprises such as *EDF* (electricity) or *SNCF* (railways)
 
**Manufacturing** industry `r fa("people-carry", fill = "green")` **unions** from the public sector `r fa("people-carry", fill = "green")` **state** policies `r fa("arrow-alt-circle-right", fill = "#cc0033")` diffusion, generalization, and **institutionalization** `r fa("arrow-alt-circle-right", fill = "#cc0033")` (almost) the whole population: directly for male wage earners and indirectly for their families.

Popular opinion: connections across related institutional domains `r fa("arrow-alt-circle-right", fill = "#cc0033")` stabilizing force.  
Reason: reformers in one area will have to consider as well the costs of possible "collateral damage" to **complementary institutions**.

**<span style="color:white;background-color:black">However</span>**, reverse effect is possible: changes in one area destabilize relations in others.   
More consistent with Germany and France: tight coupling among institutional realms `r fa("arrow-alt-circle-right", fill = "#cc0033")` reforms in adjoining policy arenas as well.
 
Germany and France: responses to the economic crisis of the 1970s and 1980s organized around saving the **core manufacturing economy**:

 - reducing the size of the workforce employed in the industrial sectors (extensive use of early retirement);
 
 - increasing the productivity among remaining workers through increased internal flexibility and intensification of work; 
 
 - negotiating new kinds of company-based deals with the remaining workforce that traded job security against increased productivity.


**<span style="color:white;background-color:black">Collateral damage</span>**: 

 - robbing the core industry of its ability to serve as lead for the rest of the economy;
 
 - new types of jobs on the outside `r fa("arrow-alt-circle-right", fill = "#cc0033")` secondary labor market for **nonstandard employment** relationships;
 
 - state institutionalization of these new types of employment: more **flexibility**, less security;
 
 - "exceptional" employment growing `r fa("arrow-alt-circle-right", fill = "#cc0033")` not allowed to compete with the core sector (i.e., putting so much
pressure on it as to compromise wages and security there);

 - growth of the secondary labor market `r fa("arrow-alt-circle-right", fill = "#cc0033")` financial and political pressure for the development of a secondary type of welfare protection;
 
 - traditionally **segregated risk pools**: 
 
   * financed by social contributions paid by employees and employers (payroll and not general taxes) 
   
   * past contributions `r fa("arrow-alt-circle-right", fill = "#cc0033")` eligibility for benefits 
   
   * increase in unemployment and in the number of jobs that are (partially) exempted from contributions 
   
   * system not designed to finance the social protection of those who did not participate in the “normal” economy and contribute to the social insurance funds
   
   * undermining financial basis of the traditional regime 
   
 - welfare reforms: sharpening the line between **social insurance** (for those who had paid their social contributions) and **social assistance** + in-work benefits for those excluded from the normal labor market and for whom the state was asked to take responsibility.
 

## Industrial Relations: Local "Egoism" and the Erosion of Collective Bargaining

<span style="color:darkred">Germany</span>:

 - **organization rate**: 35 \% in the 1970s `r fa("arrow-alt-circle-right", fill = "#cc0033")` below 25 \% today; 
 
 - **labor movement**: relatively centralized, featuring a few noncompeting multi-industrial unions;
 
 - **coordination**: manufacturing sector. 
 High levels of employer organization `r fa("people-carry", fill = "green")` pattern bargaining led informally by the powerful Metalworkers Union (*IG Metall*) `r fa("arrow-alt-circle-right", fill = "#cc0033")` 1970-80s: corporatist models like those of Scandinavia.
 
<span style="color:darkred">France</span>:

 - **organization rate**: 20 \% in 1955 `r fa("arrow-alt-circle-right", fill = "#cc0033")` < 10 \% currently;
 
 - **labor movement**: fragmented along ideological lines;
 
 - **coordination**: nationalized companies `r fa("people-carry", fill = "green")` sector- or national-level agreements `r fa("arrow-alt-circle-right", fill = "#cc0033")` all firms active within the industry or the country, even if they are signed by only a minority of the trade unions and employers’ associations

<span style="color:darkred">Both countries</span>: very high levels of **collective bargaining** coverage (at around 80 \%) `r fa("people-carry", fill = "green")` high degree of national-level harmonization of working conditions and wages.


**<span style="color:white;background-color:black">However</span>**, since 1980s: shrinkage in the traditional "core" `r fa("arrow-alt-circle-right", fill = "#cc0033")` "inward turn" on the part of firms and sectors that once led the economy.

<span style="color:darkred">Germany</span>:

 - decline of employment in manufacturing `r fa("arrow-alt-circle-right", fill = "#cc0033")` union density among the actively employed: 32.7 \% in 1980 `r fa("arrow-alt-circle-right", fill = "#cc0033")` 21.7 \% in 2004;
 
 - membership in German unions \& employer associations `r fa("arrows-alt-h", fill = "blue")` employment structures of 1960s, with strongholds in **manufacturing** and low representation in services
 
<span style="color:darkred">France</span>:

 - decline of employment in industrial sector: 25 \% in 1978 `r fa("arrow-alt-circle-right", fill = "#cc0033")` 14 \% in 2006; 
 
 - leading public companies (unions' stronghold): **privatized** from mid-1980s and increasingly after 1993;
 
 - **unionization**: 17 \% `r fa("arrow-alt-circle-right", fill = "#cc0033")` 8 \%; same composition: membership rates 3 times higher in big companies (public or private) than in small ones and 2~3 times higher in industry than in the service sectors (including services to industry)


First oil crisis `r fa("arrow-alt-circle-right", fill = "#cc0033")` protecting core skilled workers. 

"Inward turn" `r fa("arrow-alt-circle-right", fill = "#cc0033")` decentralized negotiations.

<span style="color:darkred">Germany</span>: 

 - 1972 *Revised Works Constitution Act* empowered works councils in their negotiations with management, but enhanced their autonomy vis-à-vis the overarching industrial unions;

 - 1995: position of Metal Employers’ Association (*Gesamtmetall*) collapsed when plant **managers** proved unwilling to disrupt relations with their local works councils, refusing to heed the association’s call for a lockout;
 
 - 2003: works councils in the West unwilling to continue striking for gains that would redound to the benefit of **workers** in the East. 

<span style="color:darkred">France</span>: 

 - *Auroux Laws* of 1983: government's idea to strengthen unions by anchoring their position in firms; 

 - **<span style="color:white;background-color:black">result</span>**: wounding up weakening unions by giving unorganized groups a boost;

 - wave of privatizations was explicitly relieving large, former state-owned enterprises from their public responsibilities.


**Statutory minimum wage** in Germany (&#8364; 9.82): 

 - <span style="color:darkred">supported</span> by: 
 
   * German Food and Restaurant Workers’ Union (NGG) -- representing poorly organized workers with low skills
   
   * 2004 Germany’s new (and more encompassing) *United Service Sector Union* (ver.di) 
   
 - <span style="color:darkred">opposed</span> by:
 
   * manufacturing unions -- enjoying considerable market power.
   Fearing it would compromise bargaining autonomy and put downward pressure on wages in their sectors. 
   
   * main employer confederations, the *German Employers’ Association* (BDA), and the *Confederation of German Industries* (BDI).
   
In France (&#8364; 9.67):

 - relatively high statutory wage; preventing strong wage dispersion among regular full-time employees;
 
 - previous (*Fordist*) model: every job linked to a coefficient included in salary scales; negotiated by unions and employers at the sectorial level; 
 
 - 1980s onward: **range of wages** increased as firms added different bonuses to a basic wage;
 
 - divide between workers:
 
   * **large firms**: 
   
     1) favorable deals through local bargaining for regular employees
     2) buying cheaper labor from the outside instead of hiring staff to perform services such as cleaning, security, recycling, and transportation `r fa("arrow-alt-circle-right", fill = "#cc0033")`
     3) 1959 and 2006: share of services bought by firms **multiplied by 2.5** -- fourth largest category of intermediary expenditure
   
   * **smaller firms**: outsourcing, atypical working contracts, protection and working conditions deteriorated
   
   
1993-2000 laws on working-time reduction:

 - large industrial firms and the public sector -- **unions** maintaining a presence.
 Trading working-time flexibility and increased productivity for **job security**;
 
 - small firms and low-skill service sectors: working arrangements and conditions deteriorated and **external flexibility** has increased;
 
 - high-skill and white-collar workers: working-time reduction `r fa("arrow-alt-circle-right", fill = "#cc0033")` improvements in work and life quality;
 
 - lower paid and less skilled workers: income losses, more variable schedules, and intensified work.


Germany and France -- industrial restructuring in the 1980s and 1990s:

 - reducing the size of the workforce and increasing the productivity of the remaining workers `r fa("arrow-alt-circle-right", fill = "#cc0033")` increasing competitiveness of core sectors by reducing the size of the workforce and increasing the productivity of the remaining workers;
 
 - massive use of early retirement `r fa("arrow-alt-circle-right", fill = "#cc0033")` stabilizing the remaining jobs that became harder and more flexible than before; 
 
 - rather than conflict between labor and capital, intensification of cooperation between **managers and workers** in leading firms (Germany’s manufacturing sector and in France’s large companies);
 
 - exacerbated coordination at higher levels; 
 
 - reason: unanticipated working out of developments in collective bargaining, some of which had been designed to empower organized labor; 
 
 - structures put in place in the 1970s and early 1980s to enhance labor’s voice at the plant level `r fa("arrow-alt-circle-right", fill = "#cc0033")` fueling trends toward dualism when economic hard times hit
 
 
## Labor market reforms: the institutionalization of a secondary labor market

Labor market reforms -- not affecting the status of core workers. 

<span style="color:darkred">Germany</span>: 
 
 - legal stipulations; provisions in collective contracts `r fa("arrow-alt-circle-right", fill = "#cc0033")` protecting older workers and employees with long tenure;
 
 - large and well-unionized companies: strong works councils with significant rights in the area of plant personnel policy;
 
 - mid-1980s: increase in "atypical" jobs;
 
 - late 1990s: numbers have increased steadily, even as their pay and benefits
have declined relative to "regular" employees; 

 - reforms in the 1980s and 1990s:
 
   * agency workers: 3 months in 1972 `r fa("arrow-alt-circle-right", fill = "#cc0033")` 12 months by 1997 `r fa("arrow-alt-circle-right", fill = "#cc0033")` **unlimited** in 2003; 
   
   * fixed-term contracts: 1985 \& 1996 reforms --  up to 2 years without special justification;
   
 - growth in atypical employment (primarily: mini-jobs):
 
   * previously occupied by students, housewives, and others who did not rely on such jobs as primary sources of income or benefits;
   
   * traditionally exempt from all contributions up to a threshold of earnings (in 1998 the limit was DM630); 
   
   * employees paid no payroll taxes at all;
   
   * employers could pass their taxes on to the workers, who in many cases were tax exempt;
   
   * non-wage labor costs grew in the 1980s and 1990s `r fa("arrow-alt-circle-right", fill = "#cc0033")` employers found this arrangement increasingly attractive `r fa("arrow-alt-circle-right", fill = "#cc0033")` number of such jobs increased well beyond the originally intended clientele
   
 - 1999 reform (not welcomed by employers):
 
   * replacing employer tax with employer **contributions** into the social insurance funds amounting to 22 percent of gross earnings (without, however, requiring matching contributions on the part of the worker); 
   
 - 2002 reform (welcomed by employers):
 
   * raising threshold of earnings for mini-jobs (to a maximum 400 Euro per month),
   * eliminating a previous limit of 15 hours per month,
   * allowing mini-jobs as second jobs
   
Mini-jobs: 2 million in 1991 `r fa("arrow-alt-circle-right", fill = "#cc0033")` 4.7 million in 2005. 
+ 1.7 million mini-jobs as secondary employment (thanks to the 2002 legislation).
 
**Gender component**: new possibilities for women to enter the labor market. 

**<span style="color:white;background-color:black">However</span>**: conflicts of interest among women who participate in the labor market for different reasons and on different terms:

 - single-mother mini-jobbers in Germany: 
 
   * welcoming regularization of benefits
   
   * incentives for working in mini-jobs (being contribution-free on the worker side) `r fa("arrow-alt-circle-right", fill = "#cc0033")` often yielding higher take-home pay than regular part-time work
   
   * fearing wage increases `r fa("arrow-alt-circle-right", fill = "#cc0033")` taking them above the upper earnings limit attached to mini-jobs `r fa("arrow-alt-circle-right", fill = "#cc0033")` push them into an employment zone (subject to contributions) `r fa("arrow-alt-circle-right", fill = "#cc0033")` drastically reducing their take-home pay
 
 - women who already enjoy benefits through their **insider husbands** and are simply supplementing family income: 
 
   * part of the (manufacturing-based) coalition `r fa("arrow-alt-circle-right", fill = "#cc0033")`
   * defending the status and positions of their insider husbands (mostly successfully)


<span style="color:darkred">France</span>:

 - 1988 Mitterand reelection: reinstated protections for regular workers; 
 
 - 1993: required social plans to contain detailed provisions concerning the workers to be dismissed;
 
 - shrinking core `r fa("arrow-alt-circle-right", fill = "#cc0033")` outsourcing of certain functions to smaller firms that could make use of more flexible forms of employment;
 
 - expansion of services (construction, restaurants, goods delivery, child care, elderly care) `r fa("arrow-alt-circle-right", fill = "#cc0033")` rise of a secondary labor market `r fa("arrows-alt-h", fill = "blue")` "non-standard" work contracts and lower standards (for pay, working conditions, and social protection); 
 
 - **trade unions**: successfully resisted major changes in employment protection for core workers, but experienced growing pressure to accept **increased flexibility** for other types of jobs.
 

**Atypical jobs** (fixed-term, part-time, and agency jobs): 3 \% in 1970 `r fa("arrow-alt-circle-right", fill = "#cc0033")` 25 \% by 2007. 
Currently, **70 \%** of the new job contracts are currently "atypical".

1986: enabling firms to hire temporary workers for their normal activities + extending the maximum term to 2 years.

1990-2000s: creation of subsidized, fixed-term, low-paid jobs for low-skilled workers.
Peaked in 2005 at around **500,000**.

**Official goal**: enabling the unskilled young and long-term unemployed to remain in the mainstream of society by providing a minimum income in exchange for some kind of activity, be it work or training.

1990 -- 2000: 

 - short-term employment grew by 60 \%; 
 
 - trainees, recipients of special contracts via public financing: + 65 \%; 
 
 - temporary employment grew by 130 \%;
 
 - regular employment increased by only 2 \%
 

Unions tend to support the protection of the core and flexibilization of the periphery `r fa("arrow-alt-circle-right", fill = "#cc0033")` **dual reform**

**Women**: 

 - fill 80 \% of part-time jobs;
 - 30 \% working part-time

 
## The Dualizing Dynamic of Welfare Reforms in France and Germany

1970s and 1980s: increasing numbers of inactive workers `r fa("arrow-alt-circle-right", fill = "#cc0033")` supported by fewer active workers `r fa("arrow-alt-circle-right", fill = "#cc0033")` driving up non-wage labor costs (to more than 40 \% of gross wages) and dampening job creation.

**Core problem**: growing number of workers making less contributions, drawing out of more from the system

Occupational insurance/contributory benefits `r fa("arrows-alt-h", fill = "blue")` nonoccupational/noncontributory benefits (taken care by the state)


**<span style="color:darkred">France</span>**:

1984: those with the shortest contribution records are **excluded** from any entitlement to unemployment insurance benefits `r fa("arrow-alt-circle-right", fill = "#cc0033")` pressing social problem

1988: RMI (*Revenu Minimum d’Insertion*):

 - **noncontributory** scheme for persons with no or very low income (in some cases having exhausted their right to unemployment insurance);
 - guarantees a minimum level of income-tested differential benefits to anyone aged 25 or older;
 - benefits are attached to an activation dimension, in the form of a contract between the recipient and "society";
 - recipients: obliged to take part in some **reinsertion program** (intensified searching for employment, vocational training etc.);
 - was supposed to be delivered to 300,000~400,000 people. 
 However, by March 2009, **1.13 million** persons (3 \% of the French active population) were receiving RMI;
 - currently -- **10 \%** of French population receiving welfare;
 - easing cuts in the unemployment insurance system: insurance benefits `r fa("arrow-alt-circle-right", fill = "#cc0033")` **social assistance**
 
1992: AUD (*Allocation Unique Dégressive*):

 - **deggressivity** principle -- unemployed person's insurance benefits expire entirely (after thirty months) `r fa("arrow-alt-circle-right", fill = "#cc0033")` must rely on tax-financed income-tested benefits;
 - later on (2001) -- incentives to return to the job market (negative income tax for low-paid jobs);
 - 2009 onward -- *Revenu de solidarité active* (an income given to those entering subsidized low-skill, low-paid jobs) is planned to progressively replace RMI
 

**<span style="color:darkred">Germany</sobservedpan>**:

1990s: social insurance system became untenable:

 - staggering costs of supporting the early retired and the unemployed (particularly in the East); 
 - division of labor between social insurance funds and state assistance became increasingly blurred and ultimately broke down;
 - government repeatedly called on to bail out the insurance funds on a massive scale
 
From 1998 onward:

 - Red-Green government policies similar to French;
 - **sharper line** between those supported by insurance funds + on a contributory basis and those on state-financed, income-tested assistance;
 - Harz reforms
 
**Harz IV**:

 - reducing the duration of unemployment insurance; 
   * **older** workers: 32 `r fa("arrow-alt-circle-right", fill = "#cc0033")` 18 months 
   * **other** workers: 32 `r fa("arrow-alt-circle-right", fill = "#cc0033")` 12 months
   
 - removing middle tier of unemployment assistance `r fa("arrow-alt-circle-right", fill = "#cc0033")` merging this with **social assistance**;
 - exhausting unemployment insurance benefits `r fa("arrow-alt-circle-right", fill = "#cc0033")` flat-rate income-tested social assistance;
 - *Arbeitslosengeld II* (ALG II) -- designed for those of employable age who are "able to work" and thus obliged to seek employment;
 - played into the previous labor market divisions (though not necessarily benefiting insiders -- contra Rueda): 
   * workers in **large core firms** not likely to become unemployed in the first place
   * skilled workers (particularly in the West) are highly unlikely to stay unemployed for more than a year
   * preserving a social insurance logic for skilled workers in manufacturing
   * for the less-skilled: breaking with status-protection `r fa("arrow-alt-circle-right", fill = "#cc0033")` poverty reduction
   * meeting social insurance institutions demands since the 1990s: benefit components not based on contributions out of own earnings (*versicherungsfremde Leistungen*) should not be borne by the community of insured but rather by the state
 
 - centrist unions such as the chemical workers union (representing **highly-skilled workers** with stable employment) overall **supporting** reforms;
 - **metalworkers union** -- motivated to maintain unemployment support `r fa("arrows-alt-h", fill = "blue")` "bridge" to full pensions on which early retirement agreements with employers rested `r fa("arrow-alt-circle-right", fill = "#cc0033")` **opposing** reforms;
 - **low-skill workers** and service-sector unions (*ver.di*) matter of concern: ALG II recipients have to accept any legal job they are offered, whether or not it is covered by a collective bargain;
 - contributions for regular workers (and their employers) lowered from 6.5 to 3.3 \% by 2008
 
 
**Regional dimension** (disadvantaged East vis-à-vis West):

 - long-term unemployment is a much bigger problem than in the West;
 - collective bargaining coverage (since 1980s it has included protections against layoffs of older and/or long-serving workers) is much lower than in the West;
 - protests & strikes -- mostly East participating, **no national solidarity** among unions


**<span style="color:darkred">France & Germany</span>**:

 - 1990s: changes not as a means of dismantling the Bismarckian welfare state but of **shoring it up** in a changed economic environment (slow growth, high unemployment, aging population);
 - recent reforms: strengthened the link between the amount of contribution and the volume of the benefits;
 - welfare system: relying more heavily on taxation to support the (noncontributing) working poor
 
 
Differences with a would-be typical **neoliberal deregulation**:
 
 - atypical jobs are still subject to extensive state regulation, and even outsiders benefit
from state support `r fa("arrow-alt-circle-right", fill = "#cc0033")` minimum income is guaranteed;
 - Hartz IV (the harshest of the recent reforms) -- provided a boost for an estimated one-third of those who were previously in the lowest tier of the old social assistance system;
 - **<span style="color:white;background-color:black">however</span>**, played a role in the observed increases in income inequality and poverty since the mid-1980s for Germany;
 - somewhat more egalitarian situation for France (stable inequality and poverty rates) despite weaker unions;
 - expansion of the basic safety net `r fa("arrow-alt-circle-right", fill = "#cc0033")` check on the growth of poverty in France


