---
title: "Institutionalizing Dualism"
description: |
  Lecture 3
# author:
#   - name: Kostiantyn Ovsiannikov
date: 2021-11-27
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    self_contained: false
    # code_folding: hide
creative_commons: CC BY
---


```{r setup, include=FALSE}
library(knitr)
library(here)
library(ymlthis)
library(tidyverse)
library(fontawesome)
library(readr)
library(ggthemes)
library(ggpubr)
library(RColorBrewer)
library(kableExtra)
library(data.table)
library(metathis)
library(xaringanthemer)
library(xaringanExtra)
library(data.table)
library(RefManageR)
library(plm)
library(mgcv)
library(mgcViz)
library(stargazer)
library(plotly)
library(htmlwidgets)
library(modelsummary)
```


```{r echo=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(collapse = TRUE,
                      fig.retina = 3)
```


::: {.define .article style="color: blue;"}
[Palier, B., & Thelen, K. (2010). Institutionalizing dualism: Complementarities and change in France and Germany. *Politics & Society, 38* (1), 119-148.](https://journals.sagepub.com/doi/abs/10.1177/0032329209357888)
:::

## Argument

> Contrary to theoretical accounts that suggest that institutional complementarities support stability and institutional reproduction, the authors argue that the linkages across these realms have helped to translate employer strategies that originated in the realm of industrial relations into a stable, new, and less egalitarian model with state support.


Centralized collective bargaining `r fa("arrow-alt-circle-right", fill = "#cc0033")` firm-based bargaining.

**Labor markets**: not solely deregulated, but the number of "atypical" or "nonstandard" employment relationships + working poor -- risen sharply in recent years.

Significant changes `r fa("arrows-alt-h", fill = "blue")` high degree of formal institutional stability

- negotiated and “sold” politically as ways of preserving, not undermining, traditional arrangements and the social order they reflect and represent.


**Focus of the article**: 3 institutional arenas (industrial relations, labor markets, and social protection) in the political economies of <span style="color:darkred">Germany</span> and <span style="color:darkred">France</span>.

Some authors: see it as liberalization and a move toward the American model.

This article' **argument**: politics and the outcomes `r fa("arrow-alt-circle-right", fill = "#cc0033")` institutionalization of **<span style="color:white;background-color:black">new forms of dualism</span>** (less harsh than the American model but also distinctly less egalitarian than before). 

**Cross-class coalitions** `r fa("arrow-alt-circle-right", fill = "#cc0033")` European countries not succumbing to liberalization, but also helping to promote **dualization**. 

France \& Germany: 

 - actors in the "core" economy well-positioned to defend traditional institutions and practices for themselves, but no longer serving as leaders;
 
 - able to resist outright liberalization, but especially vulnerable to dualization


More recent work on "dualization": **macro-political** perspective.   
The theory by <span style="color:blue">David Rueda</span>: 

 - labor market **insiders** and **outsiders** have distinct (and in many ways contradictory) preferences with respect to labor market policy;
 
 - insiders: preferring strong employment protections over more flexibility to ease the reentry of outsiders into the labor market; 
 
 - outsiders: preferring the opposite;

 - **social-democratic** governments `r fa("arrow-alt-circle-right", fill = "#cc0033")` interests of **insiders** over those of outsiders;
 
 - corporatist arrangements: do not stimulate social democrats to bring the interests of outsiders on board;
 
 - social democracy: not a problem (as Rueda argues) but a part of the solution 
(to resist dualism);

 - dualist trends: far more pronounced in the continental European countries than in the Nordic countries, where social democratic parties are stronger and in some cases near hegemonic;
 
 - **political coalitions** `r fa("arrow-alt-circle-right", fill = "#cc0033")` labor market policy, industrial relations, social protection.
 
 - institutionalization of dualism `r fa("arrows-alt-h", fill = "blue")` industrial relations changes, labor market policy, and welfare state reforms. 
 
Theoretical lenses: political economy `r fa("arrows-alt-h", fill = "blue")` **institutional complementarities**. 

Core industrial sectors `r fa("arrow-alt-circle-right", fill = "#cc0033")` “standard employment relationships”:

 - <span style="color:darkred">Germany</span>: large export firms in manufacturing (e.g., automobiles and machine
tools);

 - <span style="color:darkred">France</span>: nationalized firms like *Renault* + large public enterprises such as *EDF* (electricity) or *SNCF* (railways)
 
**Manufacturing** industry `r fa("people-carry", fill = "green")` **unions** from the public sector `r fa("people-carry", fill = "green")` **state** policies `r fa("arrow-alt-circle-right", fill = "#cc0033")` diffusion, generalization, and **institutionalization** `r fa("arrow-alt-circle-right", fill = "#cc0033")` (almost) the whole population: directly for male wage earners and indirectly for their families.

Popular opinion: connections across related institutional `r fa("arrow-alt-circle-right", fill = "#cc0033")` stabilizing force.  
Reason: reformers in one area will have to consider as well the costs of possible “collateral damage” to **complementary institutions**.

**<span style="color:white;background-color:black">However</span>**, reverse effect is possible: changes in one area destabilize relations in others.   
More consistent with Germany and France: tight coupling among institutional realms `r fa("arrow-alt-circle-right", fill = "#cc0033")` reforms in adjoining policy arenas as well.
 
Germany and France: responses to the economic crisis of the 1970s and 1980s organized around saving the **core manufacturing economy**:

 - reducing the size of the workforce employed in the industrial sectors (extensive use of early retirement);
 
 - increasing the productivity among remaining workers through increased internal flexibility and intensification of work; 
 
 - negotiating new kinds of company-based deals with the remaining workforce that traded job security against increased productivity.


**<span style="color:white;background-color:black">Collateral damage</span>**: 

 - robbing the core industry of its ability to serve as lead for the rest of the economy;
 
 - new types of jobs on the outside `r fa("arrow-alt-circle-right", fill = "#cc0033")` secondary labor market for **nonstandard employment** relationships;
 
 - state institutionalization of these new types of employment: more **flexibility**, less security;
 
 - "exceptional" employment growing `r fa("arrow-alt-circle-right", fill = "#cc0033")` not allowed to compete with the core sector (i.e., putting so much
pressure on it as to compromise wages and security there);

 - growth of the secondary labor market `r fa("arrow-alt-circle-right", fill = "#cc0033")` financial and political pressure for the development of a secondary type of welfare protection;
 
 - traditionally **segregated risk pools**: 
 
   * financed by social contributions paid by employees and employers (payroll and not general taxes) 
   
   * past contributions `r fa("arrow-alt-circle-right", fill = "#cc0033")` eligibility for benefits 
   
   * increase in unemployment and in the number of jobs that are (partially) exempted from contributions 
   
   * system not designed to finance the social protection of those who did not participate in the “normal” economy and contribute to the social insurance funds
   
   * undermining financial basis of the traditional regime 
   
 - welfare reforms: sharpening the line between **social insurance** (for those who had paid their social contributions) and **social assistance** + in-work benefits for those excluded from the normal labor market and for whom the state was asked to take responsibility.
