---
title: "Economic Distress, Labor Market Reforms, and Dualism in Japan and Korea"
description: |
  Lecture 5
#author: []
date: 2021-12-22
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    self_contained: false
    # code_folding: hide
creative_commons: CC BY
---

<div class="snowflakes" aria-hidden="true">
  <div class="snowflake">
  ❅
  </div>
  <div class="snowflake">
  ❅
  </div>
  <div class="snowflake">
  ❆
  </div>
  <div class="snowflake">
  ❄
  </div>
  <div class="snowflake">
  ❅
  </div>
  <div class="snowflake">
  ❆
  </div>
  <div class="snowflake">
  ❄
  </div>
  <div class="snowflake">
  ❅
  </div>
  <div class="snowflake">
  ❆
  </div>
  <div class="snowflake">
  ❄
  </div>
</div>


```{css echo = FALSE}
/*Source: https://codepen.io/codeconvey/pen/xRzQay*/
  
/* customizable snowflake styling */
.snowflake {
  color: lightblue;
  font-size: 5em;
  font-family: Arial;
  text-shadow: 0 0 1px #000;
}
@-webkit-keyframes snowflakes-fall{0%{top:-10%}100%{top:100%}}
@-webkit-keyframes snowflakes-shake{
    0%{-webkit-transform:translateX(0px);transform:translateX(0px)}
   50%{-webkit-transform:translateX(80px);transform:translateX(80px)}
  100%{-webkit-transform:translateX(0px);transform:translateX(0px)}}
@keyframes snowflakes-fall{
    0%{top:-10%}
  100%{top:100%}}
@keyframes snowflakes-shake{
    0%{transform:translateX(0px)}
   50%{transform:translateX(80px)}
  100%{transform:translateX(0px)}}
.snowflake{
  position:fixed;
  top:-10%;
  z-index:9999;
  -webkit-user-select:none;
  -moz-user-select:none;
  -ms-user-select:none;
  user-select:none;
  cursor:default;
  -webkit-animation-name:snowflakes-fall,snowflakes-shake;
  -webkit-animation-duration:10s,3s;
  -webkit-animation-timing-function:linear,ease-in-out;
  -webkit-animation-iteration-count:infinite,infinite;
  -webkit-animation-play-state:running,running;
  animation-name:snowflakes-fall,snowflakes-shake;
  animation-duration:10s,3s;
  animation-timing-function:linear,ease-in-out;
  animation-iteration-count:infinite,infinite;
  animation-play-state:running,running}
.snowflake:nth-of-type(0){left: 1%;-webkit-animation-delay:  0s,  0s;animation-delay:  0s,  0s}
.snowflake:nth-of-type(1){left:10%;-webkit-animation-delay:  1s,  1s;animation-delay:  1s,  1s}
.snowflake:nth-of-type(2){left:20%;-webkit-animation-delay:  6s,0.5s;animation-delay:  6s,0.5s}
.snowflake:nth-of-type(3){left:30%;-webkit-animation-delay:  4s,  2s;animation-delay:  4s,  2s}
.snowflake:nth-of-type(4){left:40%;-webkit-animation-delay:  2s,  2s;animation-delay:  2s,  2s}
.snowflake:nth-of-type(5){left:50%;-webkit-animation-delay:  8s,  3s;animation-delay:  8s,  3s}
.snowflake:nth-of-type(6){left:60%;-webkit-animation-delay:  6s,  2s;animation-delay:  6s,  2s}
.snowflake:nth-of-type(7){left:70%;-webkit-animation-delay:2.5s,  1s;animation-delay:2.5s,  1s}
.snowflake:nth-of-type(8){left:80%;-webkit-animation-delay:  1s,  0s;animation-delay:  1s,  0s}
.snowflake:nth-of-type(9){left:90%;-webkit-animation-delay:  3s,1.5s;animation-delay:  3s,1.5s}
```


```{r setup, include=FALSE}
library(knitr)
library(here)
library(ymlthis)
library(tidyverse)
library(fontawesome)
library(readr)
library(ggthemes)
library(ggpubr)
library(RColorBrewer)
library(kableExtra)
library(data.table)
library(metathis)
library(xaringanthemer)
library(xaringanExtra)
library(data.table)
library(RefManageR)
library(plm)
library(mgcv)
library(mgcViz)
library(stargazer)
library(plotly)
library(htmlwidgets)
library(modelsummary)
```


```{r echo=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(collapse = TRUE,
                      fig.retina = 3)
```


::: {.define .article style="color: blue;"}
[Song, J. (2012). Economic distress, labor market reforms, and dualism in Japan and Korea. *Governance, 25* (3), 415-438.](https://doi.org/10.1111/j.1468-0491.2012.01566.x)
:::

> Path-dependent trajectories of labor market and social protection reforms biased in favor of the interests of **labor market insiders** explain the rapid rise of **dualism** and inequality in the two countries.


## Intro

**VoC** framework: does not provide a sufficient explanation of the variations in labor market dualism and inequality within CMEs.

<span style="color:darkred">Japan</span>: conservative policymakers `r fa("arrows-alt-h", fill = "blue")` large firms `r fa("arrows-alt-h", fill = "blue")` core regular workers `r fa("arrow-alt-circle-right", fill = "#cc0033")` shielding labor market insiders from regulatory pressure + eroding regulations protecting outsiders for flexibility `r fa("arrow-alt-circle-right", fill = "#cc0033")` **dualism and inequality**

<span style="color:darkred">Korea</span>: strength of large **chaebol** unions `r fa("arrow-alt-circle-right", fill = "#cc0033")` obstructing the implementation of state’s labor reforms at the firm level `r fa("arrow-alt-circle-right", fill = "#cc0033")` labor market dualism and inequality


## Explaining labor market dualism and inequality in Japan and Korea

Both Japan & Korea: traditional absence of social-democratic parties `r fa("arrow-alt-circle-right", fill = "#cc0033")` weak explanatory power of "political" argument compared to EU

<span style="color:darkred">Japan</span>: 

 - liberalized temporary employment for non-regulars; 
 - decentralized wage bargaining;
 - consolidation of the dualistic welfare state

<span style="color:darkred">Korea</span>: 

 - policymakers & *chaebol* employers position: full-scale labor market reforms for insiders and outsiders + flexible labor markets;
 
 - large *chaebol* unions: fiercely opposed such reforms
 
 
## Insider-outsider differences

<span style="color:darkred">Japan's</span> labor policy: 

 - representing interests of **large firms** and their core **regular workers**;
 - deregulation of temporary employment -- **agency temps** (*haken shain*) and contract workers;
 - maintaining employment protection for insiders

Period of high growth: large Japanese firms promoted **skilled** non-regular workers to regular employment -- tight job security, wages & welfare benefits `r fa("arrow-alt-circle-right", fill = "#cc0033")` solving the problem of labor shortage

**1970s oil crises**:

 - layoffs for non-regular workers (mostly female part-time and temps workers);
 - protecting core regular male workers; 
 - share of non-regular workers -- 15-20\% till the late 1980s; 
 - most non-regulars: middle-aged **female part-timers**, secondary-income earners in families
 
 
**Worker Dispatch Law**:

 - 1986 legalized employers' rights to hire agency temps (*haken shain*) 
 - 1990 and 1996 revisions: occupational categories expanded to 16 and then to 26, but they were restricted to professionals (e.g., interpreters and software engineers);
 - 1999: "negative list system";
 - 2003: complete lifting of restrictions
 
Number of agency temps: 280,000 in 1999 `r fa("arrow-alt-circle-right", fill = "#cc0033")` 1,400,000 in 2008

**Global financial crisis**: 

 - agency temps in the manufacturing sector -- primary target of labor adjustment;
 - 207,381 workers laid off between October 2008 and June 2009;
 - \> 60\% of the laid-off workers: agency temps in the manufacturing sector
 
**In-house subcontracting**:

 - Electronics -- 53.9\%
 - Precision machinery -- 34.9\%
 
By 2010: 

 - \> 50\% of females and young workers -- non-regulars; 
 - \> 80\% of middle-aged male workers -- core regular workers

 
<span style="color:darkred">Korea</span>:

 - **1987 democratic transition** `r fa("arrow-alt-circle-right", fill = "#cc0033")` labor reform initiative;
 - chaebols & their unions: not accepting the state’s labor reform proposals for wage moderation;
 - president Kim Dae-jung in 1998: reforms to allow employers to lay off
redundant core regular workers `r fa("arrows-alt-h", fill = "blue")` improvement of workers' basic rights;
 - **chaebol unions**: blocked the initiative;
 - 2000: **Hyundai Motors** union -- pushed for regular workers' protection guarantees;
 - chaebols choice: increase in non-regular hiring + **subcontracting**;
 - SMEs' labor productivity vis-à-vis large firms: 40.3\% in 1996 `r fa("arrow-alt-circle-right", fill = "#cc0033")` 31.4% in 2004;
 - automobile recovery `r fa("cart-arrow-down", fill = "#cc0033")` no increase in the hiring of regular workers;
